Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-09-04 paper.Rnw
small updates to paper

# Specify the (base) font size for all plots
fs = 7
@

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(
    data = lime_data_gathered,
    mapping = aes(
      x = feature_stnd_value,
      y = prediction,
      size = distance_value,
      alpha = distance_value,
      color = distance
    )
  ) +
  geom_abline(
    data = coefs_data %>% filter(case == "good"),
    mapping = aes(intercept = int, slope = slope),
    size = 1
  ) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c(NA, "grey30"), guide = "none") +
  geom_point(
    data = prediction_of_interest_gathered,
    mapping = aes(x = feature_value,
                  y = prediction,
                  fill = color),
    color = "black",
    size = 5,
    shape = 23,
    alpha = 0.75
  ) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(
      feature_stnd_value = 1.2,
      prediction = 1.6,
      feature = c("Feature 1", "Feature 2"),
      slope = c(
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "good",
                   feature == "Feature 1") %>%
            pull(slope) %>%
            round(3)
        ),
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "good",
                   feature == "Feature 2") %>%
            pull(slope) %>%
            round(3)
        )
      )
    ),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = "Times",
    size = f1_fs / 2.85
  ) +
  labs(
    x = "",
    y = "Black-Box Prediction",
    title = "Conceptual Depiction of a Faithful Local Explainer Model",
    subtitle = paste("Gower Distance Metric Exponent:", good_gower_power),
    fill = "",
    alpha = "Weight",
    size = "Weight"
  ) +
  theme_linedraw(base_family = "Times", base_size = f1_fs) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f1_fs, color = "black"),
    plot.title = element_text(size = f1_fs)
  ) +
  guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2),
    alpha = guide_legend(order = 2)
  )

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(
    data = lime_data_gathered,
    mapping = aes(
      x = feature_stnd_value,
      y = prediction,
      size = distance_value,
      alpha = distance_value,
      color = distance
    )
  ) +
  geom_abline(
    data = coefs_data %>% filter(case == "bad"),
    mapping = aes(intercept = int, slope = slope),
    size = 1
  ) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c("grey30", NA), guide = "none") +
  geom_point(
    data = prediction_of_interest_gathered,
    mapping = aes(x = feature_value,
                  y = prediction,
                  fill = color),
    color = "black",
    size = 5,
    shape = 23,
    alpha = 0.75
  ) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(
      feature_stnd_value = 1.2,
      prediction = 1.6,
      feature = c("Feature 1", "Feature 2"),
      slope = c(
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "bad",
                   feature == "Feature 1") %>%
            pull(slope) %>%
            round(3)
        ),
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "bad",
                   feature == "Feature 2") %>%
            pull(slope) %>%
            round(3)
        )
      )
    ),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = "Times",
    size = f2_fs / 3
  ) +
  labs(
    x = "",
    y = "Black-Box Prediction",
    title = "Conceptual Depiction of an Unfaithful Local Explainer Model",
    subtitle = paste("Gower Distance Metric Exponent:", bad_gower_power),
    fill = "",
    alpha = "Weight",
    size = "Weight"
  ) +
  theme_linedraw(base_family = "Times", base_size = f2_fs) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f2_fs, color = "black"),
    plot.title = element_text(size = f2_fs)
  ) +
  guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2),
    alpha = guide_legend(order = 2)
  )

@@

# Plot the training data observed classes
sine_plot_obs <- ggplot(sine_data_train, aes(x = x1, y = x2, color = y)) +
  geom_point(alpha = 0.8) +
  geom_path(data = sinefun_data, aes(x = x, y = y),
            color = "black", 
            size = f3_ls) +
  scale_colour_manual(values = c(sinecolor_blue, sinecolor_red)) +
  theme_bw(base_family = "Times", base_size = f3_fs) +
  theme(aspect.ratio = 1, 
        plot.title = element_text(size = f3_fs)) +
  labs(x = TeX("$x_1$"),
       y = TeX("$x_2$"),
       color = TeX("y"),
       title = "Training Data")

# Plot the testing data rf predictions
sine_plot_pred <- ggplot() +
  geom_point(data = sine_poi %>%
               mutate(shape = "Prediction \nof Interest"),
             mapping = aes(x = x1,
                           y = x2,
                           shape = shape),
             size = 5,
             alpha = 0.8,
             color = "black") +
  geom_point(data = sine_data_test %>% filter(y != rfpred),
             mapping = aes(x = x1, y = x2),
             shape = 1,
             size = 3,
             color = "black") +
  geom_point(data = sine_data_test,
             mapping =  aes(x = x1,
                            y = x2,
                            color = rfprob_blue,
                            fill = rfprob_blue),
             shape = 21,
             alpha = 0.8) +
  geom_path(data = sinefun_data, aes(x = x, y = y),
            color = "black", 
            size = f3_ls) +
  scale_color_gradient2(low = sinecolor_red,
                        mid = '#f7f7f7',
                        high = sinecolor_blue,
                        midpoint = 0.5) +
  scale_fill_gradient2(low = sinecolor_red,
                        mid = '#f7f7f7',
                        high = sinecolor_blue,
                        midpoint = 0.5) +
  scale_shape_manual(values = 23) +
  theme_bw(base_family = "Times", base_size = f3_fs) +
  theme(aspect.ratio = 1,
        plot.title = element_text(size = f3_fs)) +
  labs(x = TeX("$x_1$"),
       y = TeX("$x_2$"),
       color = "Random \nForest \nProbability \nfor 'blue'",
       fill = "Random \nForest \nProbability \nfor 'blue'",
       title = "Testing Data",
       shape = "") +
  guides(shape = guide_legend(order = 1))

@@

# Create the lime R package explanation visualization for the sine data
sine_poi_exp <- 
  sine_poi_lime_default %>% 
  plot_features() + 
  theme(text = element_text(family = "Times", size = f4_fs),
        strip.text = element_text(size = f4_fs, face = "plain"))

# Create the explanation scatterplot for the lime explanation for the sine data
exp_scatter <-
  plot_explain_scatter(
    sine_poi_lime_default,
    alpha = 0.4,
    line_size = f4_ls
  ) +
  facet_grid(switch = "both") +
  labs(
    x = TeX("x_1"),
    y = TeX("x_2"),
    size = "Weight",
    shape = "",
    fill = "Random \nForest \nProbability \nfor 'blue'",
    color = "Random \nForest \nProbability \nfor 'blue'",
    linetype = "",
    title = "Explanation Scatterplot"
  ) +
  theme_bw(base_family = "Times", base_size = f4_fs) +
  guides(shape = guide_legend(order = 1),
         size = guide_legend(
           nrow = 2,
           byrow = T,
           override.aes = list(alpha = 1)
         )) +
  theme(aspect.ratio = 1,
        plot.title = element_text(size = f4_fs))

@@

# Create the conceptual good heatmap
heatmap_plot_good <- ggplot(data = heatmap_data_good,
                            mapping = aes(
                              x = input,
                              y = case,
                              fill = feature,
                              color = feature
                            )) +
  geom_tile() +
  labs(
    x = "Tuning Parameter Value",
    y = "Case",
    fill = "Feature",
    color = "Feature",
    title = "Situation 1",
    subtitle = "Consistent Explanations"
  ) +
  theme_bw(base_family = "Times", base_size = f5_fs) +
  scale_fill_grey() +
  scale_color_grey() +
  theme(legend.position = "none", 
        plot.title = element_text(size = f5_fs))

@@

# Create the conceptual bad heatmap
heatmap_plot_bad <- ggplot(data = heatmap_data_bad,
                           mapping = aes(
                             x = input,
                             y = case,
                             fill = feature,
                             color = feature
                           )) +
  geom_tile() +
  labs(
    x = "Tuning Parameter Value",
    y = "Case",
    fill = "Feature",
    color = "Feature",
    title = "Situation 2",
    subtitle = "Inconsistent Explanations"
  ) +
  theme_bw(base_family = "Times", base_size = f5_fs) +
  theme(plot.title = element_text(size = f5_fs)) +
  scale_fill_grey() +
  scale_color_grey()

@@

# Join the plots
plot_grid(
  heatmap_plot_good,
  heatmap_plot_bad,
  heatmap_leg,
  nrow = 1,
  rel_widths = c(0.45, 0.45, 0.1)
)

@@

# Create the example feature heatmap with the sine data
plot_feature_heatmap(sine_lime_explain$explain,
                     order_method = "PCA") +
  theme_bw(base_family = "Times", base_size = f6_fs) +
  theme(
    legend.position = "bottom",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white")
  ) +
  labs(y = "Case")

@@

# Create the metric plot
plot_metrics(sine_lime_explain$explain, rank_legend = "discrete") +
  theme_bw(base_family = "Times", base_size = f7_fs) +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.placement = "outside",
    strip.text.y = element_text(size = f7_fs)
  )

@@

<<figure-09, out.width = '3.125in', fig.width = 6, fig.height = 3, warning = FALSE>>=

@@

# Create plot of the signatures
signatures %>%
  mutate(land = c("Signature 1", "Signature 2")[as.factor(source)]) %>%
  ggplot(aes(x = x/1000)) + 
  geom_line(aes(y = sig), colour = "grey30") +
  facet_grid(land~.) +
  ylim(c(-4,6)) +
  theme_bw(base_family = "Times", base_size = f9_fs) +
  theme(strip.background = element_blank()) +
  ylab("Surface measurement (in microns)") +
  xlab("Relative Location (in millimeters)")

@@

  # Create the plot
  bullet_pcp <- bullet_train %>%
    select(all_of(bullet_features), samesource, rfscore) %>%
    bind_rows(bullet_test %>%
                select(all_of(bullet_features), samesource, rfscore),
              .id = "set") %>%
    mutate(
      set = fct_recode(set,
                       "Training Set" = "1",
                       "Testing Set" = "2"),
      samesource = fct_recode(
        factor(samesource),
        "Match" = "TRUE",
        "Non-Match" = "FALSE"
      )
    ) %>%
    rename(
      "CCF" = "ccf",
      "CMS" = "cms",
      "Matches" = "matches",
      "Mismatches" = "mismatches",
      "Non-CMS" = "non_cms",
      "Rough Correlation" = "rough_cor",
      "Distance Std Dev" = "sd_D",
      "Distance" = "D",
      "Sum Peaks" = "sum_peaks"
    ) %>%
    na.omit() %>%
    ggplot(aes(color = rfscore)) +
    geom_pcp(
      mapping = aes(vars = vars(bullet_features_ordered$feature)),
      alpha = 0.2,
      size = f10_ls
    ) +
    facet_grid(set ~ samesource) +
    scale_color_gradient2(low = "grey50",
                          high = "darkorange",
                          midpoint = 0.5) +
    theme_bw(base_family = "Times", base_size = f10_fs) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.placement = "outside",
      strip.background = element_rect(color = "white",
                                      fill = "white")
    ) +
    labs(x = "Features Ordered by Random Forest Variable Importance \n(highest to lowest from left to right)",
         y = "Scaled Feature Values",
         color = "Random \nForest \nProbability")

  # Save the plot
  ggplot2::ggsave(
    plot = bullet_pcp,
    filename = "./figure-static/figure-10.png",
    width = 7.5,
    height = 3.5,
    units = "in",
    dpi = 1200
  )

@@

  # Load the figure and then save as an eps file
  bullet_pcp <- image_read("./figure-static/figure-10-1.png")
  image_write(bullet_pcp, path = "./figure-static/figure-10-1.eps", format = "eps")

@@

# Create a feature heatmap
plot_feature_heatmap(
  bullet_explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource) %>%
    na.omit(),
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw(base_family = "Times", base_size = f11_fs) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")

@@

<<figure-12, out.width = '6.5in', warning = FALSE, fig.width = 10, fig.height = 4>>=

@@

# Create a metric comparison plot
plot_metrics(
  bullet_explain %>%
    mutate(label = as.factor(label)),
  add_lines = TRUE,
  line_alpha = 0.75,
  line_size = f12_ls
) +
  theme_bw(base_family = "Times", base_size = f12_fs) +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.placement = "outside",
    strip.text.y = element_text(size = f12_fs)
  )

@@

\begin{figure*}[!thp]
<<figure-13, out.width = '6.25in', fig.width = 13, fig.height = 17, message = FALSE>>=

@@

# Create the explanation scatterplot: case M, 3 quantile bins
eoiM_3qb <-
  plot_explain_scatter(
    bullet_explain_perms[1:3,],
    alpha = 0.9,
    weights = TRUE,
    line_size = f13_ls
  ) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = f13_fs) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = f13_fs),
    strip.text.x = element_text(size = f13_fs),
    strip.text.y = element_text(size = f13_fs),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white")
  ) +
  guides(linetype = guide_legend(override.aes = list(color = c("darkorange"))),
         size = guide_legend(nrow = 2, byrow = T))

# Create the explanation scatterplot: case MM, 3 quantile bins
eoiNM_3qb <-
  plot_explain_scatter(
    bullet_explain_perms[4:6, ],
    alpha = 0.9,
    weights = TRUE,
    line_size = f13_ls
  ) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = "Times", base_size = f13_fs) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = f13_fs),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f13_fs),
    strip.text.y = element_text(size = f13_fs)
  ) +
  guides(linetype = guide_legend(override.aes = list(color = c(
    "grey50", "darkorange"
  ))),
  size = guide_legend(nrow = 2, byrow = T))

>>>>

# Specify the (base) font size (fs) and font (ff) for all plots
fs = 7
ff = "Helvetica"
@

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(
    data = lime_data_gathered,
    mapping = aes(
      x = feature_stnd_value,
      y = prediction,
      size = distance_value,
      alpha = distance_value,
      color = distance
    )
  ) +
  geom_abline(
    data = coefs_data %>% filter(case == "good"),
    mapping = aes(intercept = int, slope = slope),
    size = 1
  ) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c(NA, "grey30"), guide = "none") +
  geom_point(
    data = prediction_of_interest_gathered,
    mapping = aes(x = feature_value,
                  y = prediction,
                  fill = color),
    color = "black",
    size = 5,
    shape = 23,
    alpha = 0.75
  ) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(
      feature_stnd_value = 1.2,
      prediction = 1.6,
      feature = c("Feature 1", "Feature 2"),
      slope = c(
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "good",
                   feature == "Feature 1") %>%
            pull(slope) %>%
            round(3)
        ),
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "good",
                   feature == "Feature 2") %>%
            pull(slope) %>%
            round(3)
        )
      )
    ),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = ff,
    size = f1_fs / 2.85
  ) +
  labs(
    x = "",
    y = "Black-Box Prediction",
    title = "Conceptual Depiction of a Faithful Local Explainer Model",
    subtitle = paste("Gower Distance Metric Exponent:", good_gower_power),
    fill = "",
    alpha = "Weight",
    size = "Weight"
  ) +
  theme_linedraw(base_family = ff, base_size = f1_fs) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f1_fs, color = "black"),
    plot.title = element_text(size = f1_fs)
  ) +
  guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2),
    alpha = guide_legend(order = 2)
  )

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(
    data = lime_data_gathered,
    mapping = aes(
      x = feature_stnd_value,
      y = prediction,
      size = distance_value,
      alpha = distance_value,
      color = distance
    )
  ) +
  geom_abline(
    data = coefs_data %>% filter(case == "bad"),
    mapping = aes(intercept = int, slope = slope),
    size = 1
  ) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c("grey30", NA), guide = "none") +
  geom_point(
    data = prediction_of_interest_gathered,
    mapping = aes(x = feature_value,
                  y = prediction,
                  fill = color),
    color = "black",
    size = 5,
    shape = 23,
    alpha = 0.75
  ) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(
      feature_stnd_value = 1.2,
      prediction = 1.6,
      feature = c("Feature 1", "Feature 2"),
      slope = c(
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "bad",
                   feature == "Feature 1") %>%
            pull(slope) %>%
            round(3)
        ),
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "bad",
                   feature == "Feature 2") %>%
            pull(slope) %>%
            round(3)
        )
      )
    ),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = ff,
    size = f2_fs / 3
  ) +
  labs(
    x = "",
    y = "Black-Box Prediction",
    title = "Conceptual Depiction of an Unfaithful Local Explainer Model",
    subtitle = paste("Gower Distance Metric Exponent:", bad_gower_power),
    fill = "",
    alpha = "Weight",
    size = "Weight"
  ) +
  theme_linedraw(base_family = ff, base_size = f2_fs) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f2_fs, color = "black"),
    plot.title = element_text(size = f2_fs)
  ) +
  guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2),
    alpha = guide_legend(order = 2)
  )

@@

# Plot the training data observed classes
sine_plot_obs <- ggplot(sine_data_train, aes(x = x1, y = x2, color = y)) +
  geom_point(alpha = 0.8) +
  geom_path(data = sinefun_data, aes(x = x, y = y),
            color = "black", 
            size = f3_ls) +
  scale_colour_manual(values = c(sinecolor_blue, sinecolor_red)) +
  theme_bw(base_family = ff, base_size = f3_fs) +
  theme(aspect.ratio = 1, 
        plot.title = element_text(size = f3_fs)) +
  labs(x = TeX("$x_1$"),
       y = TeX("$x_2$"),
       color = TeX("y"),
       title = "Training Data")

# Plot the testing data rf predictions
sine_plot_pred <- ggplot() +
  geom_point(data = sine_poi %>%
               mutate(shape = "Prediction \nof Interest"),
             mapping = aes(x = x1,
                           y = x2,
                           shape = shape),
             size = 5,
             alpha = 0.8,
             color = "black") +
  geom_point(data = sine_data_test %>% filter(y != rfpred),
             mapping = aes(x = x1, y = x2),
             shape = 1,
             size = 3,
             color = "black") +
  geom_point(data = sine_data_test,
             mapping =  aes(x = x1,
                            y = x2,
                            color = rfprob_blue,
                            fill = rfprob_blue),
             shape = 21,
             alpha = 0.8) +
  geom_path(data = sinefun_data, aes(x = x, y = y),
            color = "black", 
            size = f3_ls) +
  scale_color_gradient2(low = sinecolor_red,
                        mid = '#f7f7f7',
                        high = sinecolor_blue,
                        midpoint = 0.5) +
  scale_fill_gradient2(low = sinecolor_red,
                        mid = '#f7f7f7',
                        high = sinecolor_blue,
                        midpoint = 0.5) +
  scale_shape_manual(values = 23) +
  theme_bw(base_family = ff, base_size = f3_fs) +
  theme(aspect.ratio = 1,
        plot.title = element_text(size = f3_fs)) +
  labs(x = TeX("$x_1$"),
       y = TeX("$x_2$"),
       color = "Random \nForest \nProbability \nfor 'blue'",
       fill = "Random \nForest \nProbability \nfor 'blue'",
       title = "Testing Data",
       shape = "") +
  guides(shape = guide_legend(order = 1))

@@

# Create the lime R package explanation visualization for the sine data
sine_poi_exp <- 
  sine_poi_lime_default %>% 
  plot_features() + 
  theme(text = element_text(family = ff, size = f4_fs),
        strip.text = element_text(size = f4_fs, face = "plain"))

# Create the explanation scatterplot for the lime explanation for the sine data
exp_scatter <-
  plot_explain_scatter(
    sine_poi_lime_default,
    alpha = 0.4,
    line_size = f4_ls
  ) +
  facet_grid(switch = "both") +
  labs(
    x = TeX("x_1"),
    y = TeX("x_2"),
    size = "Weight",
    shape = "",
    fill = "Random \nForest \nProbability \nfor 'blue'",
    color = "Random \nForest \nProbability \nfor 'blue'",
    linetype = "",
    title = "Explanation Scatterplot"
  ) +
  theme_bw(base_family = ff, base_size = f4_fs) +
  guides(shape = guide_legend(order = 1),
         size = guide_legend(
           nrow = 2,
           byrow = T,
           override.aes = list(alpha = 1)
         )) +
  theme(aspect.ratio = 1,
        plot.title = element_text(size = f4_fs))

@@

# Create the conceptual good heatmap
heatmap_plot_good <- ggplot(data = heatmap_data_good,
                            mapping = aes(
                              x = input,
                              y = case,
                              fill = feature,
                              color = feature
                            )) +
  geom_tile() +
  labs(
    x = "Tuning Parameter Value",
    y = "Case",
    fill = "Feature",
    color = "Feature",
    title = "Situation 1",
    subtitle = "Consistent Explanations"
  ) +
  theme_bw(base_family = ff, base_size = f5_fs) +
  scale_fill_grey() +
  scale_color_grey() +
  theme(legend.position = "none", 
        plot.title = element_text(size = f5_fs))

@@

# Create the conceptual bad heat map
heatmap_plot_bad <- ggplot(data = heatmap_data_bad,
                           mapping = aes(
                             x = input,
                             y = case,
                             fill = feature,
                             color = feature
                           )) +
  geom_tile() +
  labs(
    x = "Tuning Parameter Value",
    y = "Case",
    fill = "Feature",
    color = "Feature",
    title = "Situation 2",
    subtitle = "Inconsistent Explanations"
  ) +
  theme_bw(base_family = ff, base_size = f5_fs) +
  theme(plot.title = element_text(size = f5_fs)) +
  scale_fill_grey() +
  scale_color_grey()

@@

# Join the plots
plot_grid(
  heatmap_plot_good,
  heatmap_plot_bad,
  heatmap_leg,
  nrow = 1,
  rel_widths = c(0.43, 0.43, 0.14)
)

@@

# Create the example feature heatmap with the sine data
plot_feature_heatmap(sine_lime_explain$explain,
                     order_method = "PCA") +
  theme_bw(base_family = ff, base_size = f6_fs) +
  theme(
    legend.position = "bottom",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white")
  ) +
  labs(y = "Case")

@@

# Create the metric plot
plot_metrics(sine_lime_explain$explain, rank_legend = "discrete") +
  theme_bw(base_family = ff, base_size = f7_fs) +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.placement = "outside",
    strip.text.y = element_text(size = f7_fs)
  )

@@

<<figure-09, out.width = '3.125in', fig.width = 6, fig.height = 3.5, warning = FALSE>>=

@@

# Create plot of the signatures
signatures %>%
  mutate(land = c("Signature 1", "Signature 2")[as.factor(source)]) %>%
  ggplot(aes(x = x/1000)) + 
  geom_line(aes(y = sig), colour = "grey30") +
  facet_grid(land~.) +
  ylim(c(-4,6)) +
  theme_bw(base_family = ff, base_size = f9_fs) +
  theme(strip.background = element_blank()) +
  ylab("Surface measurement (in microns)") +
  xlab("Relative Location (in millimeters)")

@@

  # Create the plot
  bullet_pcp <- bullet_train %>%
    select(all_of(bullet_features), samesource, rfscore) %>%
    bind_rows(bullet_test %>%
                select(all_of(bullet_features), samesource, rfscore),
              .id = "set") %>%
    mutate(
      set = fct_recode(set,
                       "Training Set" = "1",
                       "Testing Set" = "2"),
      samesource = fct_recode(
        factor(samesource),
        "Match" = "TRUE",
        "Non-Match" = "FALSE"
      )
    ) %>%
    rename(
      "CCF" = "ccf",
      "CMS" = "cms",
      "Matches" = "matches",
      "Mismatches" = "mismatches",
      "Non-CMS" = "non_cms",
      "Rough Correlation" = "rough_cor",
      "Distance Std Dev" = "sd_D",
      "Distance" = "D",
      "Sum Peaks" = "sum_peaks"
    ) %>%
    na.omit() %>%
    ggplot(aes(color = rfscore)) +
    geom_pcp(
      mapping = aes(vars = vars(bullet_features_ordered$feature)),
      alpha = 0.2,
      size = f10_ls
    ) +
    facet_grid(set ~ samesource) +
    scale_color_gradient2(low = "grey50",
                          high = "darkorange",
                          midpoint = 0.5) +
    theme_bw(base_family = ff, base_size = f10_fs) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.placement = "outside",
      strip.background = element_rect(color = "white",
                                      fill = "white")
    ) +
    labs(x = "Features Ordered by Random Forest Variable Importance \n(highest to lowest from left to right)",
         y = "Scaled Feature Values",
         color = "Random \nForest \nProbability")

  # Save the plot
  ggplot2::ggsave(
    plot = bullet_pcp,
    filename = "./figure-static/figure-10-1.png",
    width = 7.5,
    height = 3.5,
    units = "in",
    dpi = 1200
  )

@@

  # Load the figure and then save as an EPS file
  bullet_pcp <- image_read("./figure-static/figure-10-1.png")
  image_write(bullet_pcp, path = "./figure-static/figure-10-1.eps", format = "eps")

@@

# Create a feature heatmap
plot_feature_heatmap(
  bullet_explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource) %>%
    na.omit(),
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw(base_family = ff, base_size = f11_fs) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")

@@

<<figure-12, out.width = '6.5in', warning = FALSE, fig.width = 10, fig.height = 4.25>>=

@@

# Create a metric comparison plot
plot_metrics(
  bullet_explain %>%
    mutate(label = as.factor(label)),
  add_lines = TRUE,
  line_alpha = 0.75,
  line_size = f12_ls
) +
  theme_bw(base_family = ff, base_size = f12_fs) +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.placement = "outside",
    strip.text.y = element_text(size = f12_fs)
  )

@@

\begin{figure*}[!thp]
<<figure-13, out.width = '6.25in', fig.width = 13, fig.height = 17.5, message = FALSE>>=

@@

# Create the explanation scatterplot: case M, 3 quantile bins
eoiM_3qb <-
  plot_explain_scatter(
    bullet_explain_perms[1:3,],
    alpha = 0.9,
    weights = TRUE,
    line_size = f13_ls
  ) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = ff, base_size = f13_fs) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = f13_fs),
    strip.text.x = element_text(size = f13_fs),
    strip.text.y = element_text(size = f13_fs),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white")
  ) +
  guides(linetype = guide_legend(override.aes = list(color = c("darkorange"))),
         size = guide_legend(nrow = 2, byrow = T))

# Create the explanation scatterplot: case MM, 3 quantile bins
eoiNM_3qb <-
  plot_explain_scatter(
    bullet_explain_perms[4:6, ],
    alpha = 0.9,
    weights = TRUE,
    line_size = f13_ls
  ) +
  scale_color_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  scale_fill_gradient2(
    low = "grey50",
    high = "darkorange",
    midpoint = 0.5,
    limits = c(0, 1)
  ) +
  theme_bw(base_family = ff, base_size = f13_fs) +
  theme(
    aspect.ratio = 1,
    plot.title = element_text(size = f13_fs),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f13_fs),
    strip.text.y = element_text(size = f13_fs)
  ) +
  guides(linetype = guide_legend(override.aes = list(color = c(
    "grey50", "darkorange"
  ))),
  size = guide_legend(nrow = 2, byrow = T))

### 2020-09-04 paper.tex
small updates to paper

%The following supporting information is available as part of the online article:

>>>>

The following supporting information is available:

\kge{\begin{itemize}
\item \textbf{Paper GitHub Repository}: The R markdown document and code used to produce the paper are available in the following GitHub Repository: \href{https://github.com/goodekat/LIME-diagnostics-paper-short-sweet}{https://github.com/goodekat/LIME-diagnostics-paper-short-sweet}. \kgc{Need to update this link after I change the name of the repository.}
\item \textbf{Data}: The data used for the examples in Section \ref{application} are available at \kgc{need to fill in once Heike has the data published.}
\end{itemize}}
